# Lecture 6 -- Confidence Intervals {.unnumbered}

## Guided Notes

FILES: BLANK [lecture-6.pdf](https://github.com/coltongearhart/math321/blob/main/lectures/lecture-6.pdf), [lecture-6.tex](https://github.com/coltongearhart/math321/blob/main/lectures/lecture-6.tex), and [style-lectures.sty](https://github.com/coltongearhart/math321/blob/main/lectures/style-lectures.sty)

<embed src="lectures/lecture-6-COMPLETED.pdf" type="application/pdf" width="100%" height="1000px"></embed>

## In-Class Activity

FILES: [in-class-6.tex](https://github.com/coltongearhart/math321/blob/main/assessments/in-class-6.tex) and [style-assessments.sty](https://github.com/coltongearhart/math321/blob/main/assessments/style-assessments.sty)

Instructors can contact me for keys :)

<embed src="assessments/in-class-6.pdf" type="application/pdf" width="100%" height="1000px"></embed>

## Homework

FILES: Excel [homework-6-data.xlsx](https://github.com/coltongearhart/math321/blob/main/excel/homework-6-data.xlsx), R [homework-6-data.RData](https://github.com/coltongearhart/math321/blob/main/r/files/data/homework-6-data.RData),  [homework-6.tex](https://github.com/coltongearhart/math321/blob/main/assessments/homework-6.tex) and [style-assessments.sty](https://github.com/coltongearhart/math321/blob/main/assessments/style-assessments.sty)

- If doing the assignment in R: After saving the .RData file in the same directory as your R script, run the following command to load the data.

```{r}

# load data
load("r/files/data/homework-6-data.RData")

```

```{r}
#| echo: false

# preview datasets
for (obj in ls()) {
  cat('---', obj, '---\n') # print variable name with formatting
  str(get(obj)) # display structure
}

```

Instructors can contact me for solutions :)

<embed src="assessments/homework-6.pdf" type="application/pdf" width="100%" height="1000px"></embed>