\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)
\newcommand{\blankul}[1]{\rule[-1.5mm]{#1}{0.15mm}}	% shortcut for blank underline, where the only option needed to specify is length (# and units (cm or mm, etc.)))

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: Homework 10}
\end{center}

\bigskip\bigskip

{\large \textbf{Due \blankul{4cm}}: Turn in a hard copy, neat and stapled.}\bigskip

% problem types summary
% 1) Binomial mean and var derivation
% 2) Discrete uniform mean and sd
% 3) Binomial probability used in a tree diagram
% 4) Negative binomial probabilities
% 5) Mean of a function of a negative binomial
% 6) Hypergeometric probabilities and exp value and sd
% 7) Poisson probabilities (compound events)
% 8) Poisson pmf to find variance

% a good Poisson problem is Lazar HW 4 Additional Questions #2

\begin{enumerate}
    \item For a binomial random variable $X$ with $n = 2$ and $P(\text{Success}) = p$, show using the definitions that% Actex 5-10
    \begin{enumerate}
        \item $E(X) = 2p$ \hspace{85pt} (i.e. use $E(X) = \sum x f(x))$
        \item $V(X) = 2p(1 - p) = 2pq $\hspace{30pt} (You may use the alternate form of variance)
    \end{enumerate}\bigskip
    
    \item A contestant on a game show selects a ball from a basket containing 25 balls numbered from 1 to 25. Their prize is \$850 times the number of the ball selected. Find the mean and standard deviation of the amount they win.\bigskip%Actex 5-34 (with different numbers)
    
    \item A nutrition company receives 2/5 of its supplement shipments from company $X$ and the remainder of its shipments from other companies. Each shipment contains a very large number of supplement bottles.% Actex 5-38 (with different numbers and context)
    \item[] For Company $X$'s shipments, 7\% of the bottles are mislabelled. For every other company, 12\% of the bottles are mislabelled.
    \item[] The nutrition company inspects 25 randomly selected bottles from a single shipment and finds that one bottle is mislabelled. Find the probability that the shipment came from Company $X$.
    
    \item A telemarketer makes successful calls with probability 0.27. Their shift ends when they make 4 sales. Find the following:%Actex 5-30 (with different numbers and slightly different wording)
    \begin{enumerate}
        \item The probability that the 4\textsuperscript{th} sale will be on the 13\textsuperscript{th} call.
        \item The probability it will take more than 5 calls to make the 4 sales.
    \end{enumerate}\bigskip
    
    \item Suppose now that each sale made by the person in problem 4 is for \$300. Find the mean number of total calls they will have to make to reach \$2400.\bigskip%Actex 5-31 (with different numbers)
    
    \item In a shipment of lightbulbs 120 lightbulbs, there are 12 defective ones. An inspector randomly selects 15 bulbs. Let $X$ represent the number of defective light bulbs selected by the inspector.%Prob and stat inference 2.5-1 (for the context)
    \begin{enumerate}
        \item Find the pmf of $X$.
        \item Find the probability less than 3 defective light bulbs are found. 
        \item Find the probability at least one defective light bulb is found.
        \item Find the expected value and standard deviation of the number of defective light bulbs found.
    \end{enumerate}\bigskip
    
    \newpage
    
    \item Claims filed in a year by a policyholder of an insurance company have a Poisson distribution with $\lambda = 0.6$. The number of claims filed by two different policyholders are independent events.%Actex 5-20 (with different numbers)
    \begin{enumerate}
        \item If two policyholders are selected at random, find the probability that each of them will file one claim during the year.
        \item Find the probability that at least one of them will file zero claims.
    \end{enumerate}\bigskip
    
    \item An actuary has discovered that policyholders are four times as likely to file two claims as to file three claims. If the number of claims filed as a Poisson distribution, find the variance of the number of claims filed.%Actex 5-39 (with different numbers)
    
\end{enumerate}

\vspace{100pt}

Select answers\bigskip
\begin{enumerate}
    \item 
    \begin{enumerate}
        \item 
        \item 
    \end{enumerate}
    
    \item $SD(\text{Money}) \approx \$6,129.44$
    
    \item $\text{Prob} \approx 0.5943$
    
    \item 
    \begin{enumerate}
        \item $\text{Prob} \approx 0.0688$
        \item $\text{Prob} \approx 0.9792$
    \end{enumerate}
    
    \item $\text{Exp Value} \approx 29.63$
    
    \item 
    \begin{enumerate}
        \item $\text{Prob} \approx 0.8269$
        \item $\text{Prob} \approx 0.8149$
        \item $SD \approx 1.0914$
        \item 
    \end{enumerate}
    
    \item 
    \begin{enumerate}
        \item $\text{Prob} \approx 0.1084$
        \item $\text{Prob} \approx 0.7964$
    \end{enumerate}
    
    \item 
      
\end{enumerate}  
    
\end{document}