\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)
\newcommand{\blankul}[1]{\rule[-1.5mm]{#1}{0.15mm}}	% shortcut for blank underline, where the only option needed to specify is length (# and units (cm or mm, etc.)))
\newcommand{\follow}[1]{\sim \text{#1}\,}		% shortcut for ~ 'Named dist ' in normal font with space before parameters would go
\newcommand{\integral}[4]{\displaystyle \int_{#1}^{#2} #3 \,\mathrm{d} #4}		% shortcut for large integral with limits and appending formatted dx (variable x)
\newcommand{\e}{\mathrm{e}}		% shortcut for non-italic e in math mode

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: Homework 11}
\end{center}

\bigskip\bigskip

{\large \textbf{Due \blankul{4cm}}: Turn in a hard copy, neat and stapled.}\bigskip

% problem types summary
% 1) Uniform (probabilities, expected value, sd and percentiles)
% 2) Derivation of exponential mean
% 3) Exponential probabilities and expected value
% 4) Derivation of exponential median
% 5) Exponential probabilities (including memoryless)
% 6) Exponential probabilities and gamma mean / variance
% 7) Gamma mean and variance
% 8) Gamma probability (integration by parts practice)
% 9) CLT based on uniform RV
% 10) CLT percentile
% 11) CLT based on exponential RV

\begin{enumerate}
    \item A professor gives an exam where the time limit is 60 min and the anticipated minimum time to students to take the exam is 35 min. Assume that the random variable $T$ for the time it takes a student to finish the exam is uniformly distributed over $[35,60]$.%Actex 8-5 (with different numbers)
    \begin{enumerate}
        \item Find the probability that it takes a student between 40 and 55 min to finish the exam.%Original addition
        \item Find the $E(T)$ and $SD(T)$.
        \item Find the time $T$ that 75\% of students will be finished with the exam.
        \item Suppose you and your friend are taking the exam (independently). Find the probability that both of you are finished within 50 min.%Original addition
    \end{enumerate}\bigskip
    
    \item Using the definition of expected value $E(X) = \integral{-\infty}{\infty}{x f(x)}{x}$, show that if $T \follow{Exponential}(\lambda)$, then $E(T) = 1 / \lambda$.\bigskip%Original
    
    \item 
    \begin{enumerate}
        \item Let $T \follow{Exponential}(\lambda = 5)$. Find $P[T < E(T)]$.
        \item Let $T \follow{Exponential}(\lambda = 0.15)$. Find $P[T < E(T)]$.
        \item Let $T \follow{Exponential}(\lambda)$. Find $P[T < E(T)]$.
    \end{enumerate}\bigskip
    
    \item If $T \follow{Exponential}(\lambda)$, find the median of $T$.\bigskip%Actex 8-10
    \item Researchers at a medical facility have discovered a virus whose mean time from initial infection to symptoms appearing is 30 days. Assume this time $T$ has an exponential distribution.%Actex 8-13 (with different numbers)
    \item[] \textit{HINT: Be careful with the parameter.}
    \begin{enumerate}
        \item Find the probability that a patient who has just been infected will show symptoms in 25 days.
        \item Find the probability that a patient who has just been infected will not show symptoms for at least 33 days.
        \item Given that an infected patient has been without symptoms for the first 14 days, find the probability they will not show symptoms for at least the next 7 days.%Original
    \end{enumerate}\bigskip
    
    \item Suppose hurricanes in Florida occur following a Poisson process at a rate of 0.02 per year, and the waiting time $T$ between hurricanes can be modeled with an exponential random variable.%Actex 8-22 (setup, different context)
    \begin{enumerate}
        \item Find the probability that the first hurricane happens before year 5.%Original
        \item Find the probability that the first hurricane happens between years 7 and 10.%Original
        \item Let $W$ be the waiting time from the start of observation until the third hurricane. Find $E(W)$ and $V(W)$.%Actex 8-22
    \end{enumerate}\bigskip
    
    \item A gamma distribution has a mean of 9 and a variance of 16. Find $\alpha$ and $\beta$ for this distribution.\bigskip%Actex 8-24 (with different numbers)
    
    \item Let $X \follow{Gamma}(\alpha = 2, \beta = 1/2)$. Find $P(X < 3)$.\bigskip%Original
    
    \item If a number is selected at random from the interval $[0,100]$, its value has a uniform distribution over that interval. Let $S$ be the random random variable for the sum of 45 numbers selected at random from $[0,100]$. Find $P(2050 < S < 2650)$.\bigskip%Actex 8-34 (with different numbers)
    
    \item A charity receives 2000 contributions. Contributions are assumed to be independent and identically distributed with mean \$3500 and standard deviation \$300. Calculate the approximate $95^{th}$ percentile for this distribution for the total contributions received.\bigskip%Actex 8-66 (with different numbers)
    
    \item The total claim amount for a health insurance policy follows a distribution with density function%Actex 8-63 (with different numbers)
    \[f(x) = \frac{1}{700} \e^{-x / 700} \quad\quad x \ge 0.\]
    \item[] The premium (cost of the policy for policyholders) is set at 50 over the expected total claim amount.
    \item[] If 100 policies are sold, find the approximate probability that the insurance company will have claims (total losses on policies) exceeding the premiums collected (total money earned from policyholder payments). Assume all policies are independent and have identical claim amount distributions.
\end{enumerate}

\bigskip

Select answers\bigskip
\begin{enumerate}
    \item 
    \begin{enumerate}
        \item $\text{Prob} = 0.6$
        \item 
        \item 53.75 min
        \item $\text{Prob} = 0.36$
    \end{enumerate}
    
    \item 
    
    \item 
    \begin{enumerate}
        \item 
        \item 
        \item $\text{Prob} \approx 0.632$
    \end{enumerate}
      
    \item $m \approx 0.693 / \lambda$
    
    \item  
    \begin{enumerate}
        \item $\text{Prob} \approx 0.565$
        \item $\text{Prob} \approx 0.333$
        \item $\text{Prob} \approx 0.7918$
    \end{enumerate}
    
    \item  
    \begin{enumerate}
        \item $\text{Prob} \approx 0.095$
        \item $\text{Prob} \approx 0.0506$
        \item $E(W) = 150$
    \end{enumerate}
    
    \item $\alpha \approx 5.0625$ and $\beta \approx 0.5625$
    
    \item $\text{Prob} \approx 0.442$
    
    \item $\text{Prob} \approx 0.8297$ using calc or $\text{Prob} \approx 0.8293$ using table

    \item $\approx 7,022,068.03$ using calc or 7,022,137.03 using table
    
    \item $\text{Prob} \approx 0.2375$ using calc or $\text{Prob} \approx 0.2389$ using table


\end{enumerate}
    
\end{document}