\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)
\newcommand{\blankul}[1]{\rule[-1.5mm]{#1}{0.15mm}}	% shortcut for blank underline, where the only option needed to specify is length (# and units (cm or mm, etc.)))
\newcommand{\comp}{{\sim}}						% shortcut for tilde without extra space, using for complement

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: Homework 4-5}
\end{center}

\bigskip\bigskip

{\large \textbf{Due \blankul{4cm}}: Turn in a hard copy, neat and stapled.}\bigskip

% problem types summary
% 1) Updating scenario conditional probability
% 2) Conditional probability with Venn diagram
% 3) Independence proof
% 4) Checking independence
% 5) Using independence multiplication rule to solve for unknown info
% 6) Using independence in calculations


\begin{enumerate}
    \item A swim team consists of 6 boys and 4 girls. A relay team of 4 swimmers is chosen at random from the team members. What is the probability that 2 boys are selected for the relay team given that the first 2 selections were girls?\bigskip% MATH 125 Module 7.3 question
    
    \item An actuary is studying the prevalence of three health risk factors, denoted A, B and C, within a population of women. For each of the three factors, the probability is 0.12 that a women in the population has only this risk factor (and no others). For any two of the three risk factors, the probability is 0.15 that she has exactly these two risk factors (but not the other). The probability that a woman has all three risk factors, given that she has A and B, is 1/3.\bigskip%Actex 3-56 (with different numbers, and broken up questions)
    \begin{enumerate}
        \item Find the probability that a woman has all three risk factors.
        \item Find the probability that a woman has none of the three risk factors, given that she does not have risk factor A.
    \end{enumerate}\bigskip

    \item If $A$ and $B$ are independent events, prove $\comp A$ and $\comp B$ are also independent.\bigskip%Prob and stat inference 1.4-4
    
    \item For the experiment of tossing a single fair coin 3 times, let $E$ be the event the first toss is a head and $F$ be the event 2 heads and 1 tail are tossed. Show if $E$ and $F$ are independent.\bigskip%Actex 3-37
    
    \item A jar contains 10 marbles: 4 red and 6 blue. A second jar contains 16 red marbles and an unknown number of blue marbles. A single marble is drawn from each jar. % Actex 3-55 
   \begin{enumerate}
        \item Suppose the probability that both marbles are red is 0.256. Calculate the number of blue marbles in the second jar.% Original addition to ease into it
        \item Now suppose the probability that both marbles are the same color is 0.44. Calculate the number of blue marbles in the second jar.
    \end{enumerate}
    
    \item An insurance company pays hospital claims. The number of claims that include emergency room or operating room charges is 85\% of the total number of claims. The number of claims that do not include emergency room charges is 25\% of the total number of claims. The occurrence of emergency room charges is independent of the occurrence of operating room charges on hospital claims.%Actex 3-52
    \item[] Calculate the probability that a claim submitted to the insurance company includes operating room charges.\bigskip
    
\end{enumerate}

\newpage

\begin{enumerate}
    \item $\text{Prob} \approx 0.536$
    
    \item 
    \begin{enumerate}
        \item $\text{Prob} = 0.075$
        \item $\text{Prob} \approx 0.228$
    \end{enumerate}
    
    \item 
    
    \item 
    
    \item 
    \begin{enumerate}
        \item 4
        \item 9
    \end{enumerate}
    
    \item $\text{Prob} = 0.4$
\end{enumerate}

\end{document}