\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: In-Class 4}

\end{center}

\bigskip\bigskip


Answer all questions. Show your work where necessary.\bigskip

% problem types summary
% 1) Direct counting problem
% 2) Conditional probability by counting outcomes
% 3) Conditional probability definition
% 4) Conditional probability by counting outcomes and then with a tree
% 5) Conditional probability with random variable
% 6) Monty Hall problem

\begin{enumerate}
    \item Four cards are to be dealt successively at random and without replacement from an ordinary deck of playing cards. What is the probability of receiving in order a spade, heart, spade, and a club?\vspace{40pt}% Lazar Problem 1.3-5
    
    \item In an experiment of tossing a single fair coin three times, what is the probability of getting exactly 2 tails, given that you get at least one tails?\vspace{100pt}%Actex 3-27
    \item If two events, $A$ and $B$, are such that $P(A) = 0.4$, $P(B) = 0.25$ and $P(A \cap B) = 0.1$, find the following:%Math stats with apps 2-71 (with different numbers)
    \begin{enumerate}
        \item $P(A \mid B)$\vspace{30pt}
        \item $P(B \mid A)$\vspace{30pt}
        \item $P(A \mid A \cup B)$\vspace{30pt}
        \item $P(A \mid A \cap B)$\vspace{30pt}
        \item $P(A \cap B \mid A \cup B)$\vspace{30pt}
    \end{enumerate}\bigskip
    
    \item 
    \begin{enumerate}
        \item A woman has two children. She tells you that at least one of them is a boy. What is the probability that both children are boys? Assume $P(\text{Boy}) = 0.5$.\vspace{50pt}% Lazar Problem 1.3-7
        \item Same question as part (a), but now assume $P(\text{Boy}) = 0.47$.\vspace{70pt}
    \end{enumerate}

    \item The number of injury claims per month is modeled by a random variable $N$ with \\$P(N = n) = \frac{1}{(n + 1)(n + 2)}$, where $n \ge 0$.%Actex 3-53 (with different question)
    \item[] Determine the probability of exactly one claim during the particular month, given there have been at most two claims.\vspace{100pt}
    
    \item Monty Hall problem: In the game show ``Let's Make a Deal'', a contestant is presented with 3 doors. There is a prize behind one of the doors, and the host of the show knows which one. When the contestant makes a choice of door, at least one of the other doors will not have a prize, and the host will open a door (one not chosen by the contestant) with no prize. The contestant is given the option to change his choice after the host shows the door without a prize.% Lazar Problem 1.3-9
    \begin{enumerate}
    \item If the contestant switches doors, what is the probability that they get the door with the prize?\vspace{100pt}
    \item If the contestant does NOT switch doors, what is the probability that they get the door with the prize?\vspace{100pt}
    \end{enumerate}
    

\end{enumerate}

\end{document}