\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: In-Class 10}

\end{center}
\bigskip\bigskip

Assume that 18\% of people are left handed. Answer the following questions based on each described experiment.\bigskip
\begin{enumerate}
    \item Suppose we select 9 people at random. Let $X$ be the number of lefties selected out of the 9 people.\bigskip
    \begin{enumerate}
        \item Find the probability there are exactly 4 lefties.\vspace{80pt}%binomial P(X = 4)
        \item Find the probability there are at least 6 lefties.\vspace{80pt}%binomial P(X >= 6)
        \item Find the probability there are is a majority of righties.\vspace{80pt}%binomial P(X <= 4) OR P(Y >= 5)
        \item How many lefties do you expect in the group? With what standard deviation?\vspace{80pt}%binomial E(X) and SD(X)
        \item Suppose there are less than 4 lefties in the group. Find the probability there are exactly 2 lefties?\vspace{80pt}%binomial P(X = 2 | X < 4)
    \end{enumerate}\newpage
    
    \item Suppose we select people at random until the first lefty is selected. Let $Y$ be the number of people selected in order to select the first lefty.\bigskip
    \begin{enumerate}
        \item Find the probability the first lefty is the 5\textsuperscript{th} person.\vspace{80pt}%geometric P(X = 5)
        \item Find the probability the first lefty is before the 7\textsuperscript{th} person.\vspace{80pt}%geometric P(X < 7)
        \item Find the probability the first lefty is the 3\textsuperscript{rd} through  6\textsuperscript{th} person.\vspace{80pt}%geometric P(3 <= X <= 6)
        \item How many people do you expect to select until the first lefty? With what standard deviation?\vspace{80pt}%geometric E(X) and SD(X)
        \item Suppose the first 5 people were righties, find the probability the first 9 people are righties.\vspace{80pt}%geometric P(X > 9 | X > 5)
    \end{enumerate}
\end{enumerate}

\end{document}