\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: In-Class 6}

\end{center}

\bigskip\bigskip


Answer all questions. Show your work where necessary.\bigskip

% problem types summary
% 1) Law of total probability and Bayes' theorem (tree diagram)
% 2) Law of total probability and Bayes' Theorem (tree diagram), and using complements
% 3) Law of total probability
% 4) Bayes' Theorem
% 5) Law of total probability and Bayes' theorem (harder figuring out stages)
% 4) Bayes' Theorem (harder, unknown conditional probability x)

\begin{enumerate}
    
    \item The manufacturing company has a fabrication plant and an assembly line. The fabrication plant has 65\% of the employees and the assembly line has 35\%. During the past year 25\% of the workers in the fabrication plant sustained injuries and 15\% of the assembly line workers had injuries.%Actex 3-28 (with different numbers)
    \begin{enumerate}
        \item What percentage of all workers had injuries in this period? Draw a tree diagram to help.\vspace{100pt}
        \item If an employee had an injury, what is the probability that they worked on the assembly line?\vspace{40pt}
    \end{enumerate}\bigskip
    
    \item An insurance company estimates that 40\% of policyholders who have only an auto policy will renew next year and 60\% of policyholders who have only a homeowners policy will renew next year. The company estimates that 80\% of policyholders who have both an auto and a homeowners policy will renew next year. Company records show that 65\% of policyholders have an auto policy, 50\% of policyholders have a homeowners policy and 15\% of policyholders have both an auto and a homeowners policy.% Lazar 1.5-3 (with additions)
    \begin{enumerate}
        \item What percentage of policyholders will renew at least one policy next year?\vspace{100pt}
        \item What percentage of policyholders will not renew at least one policy next year?\vspace{40pt}
        \item Given that a customer renews, what is the probability they have only an auto policy?\vspace{40pt}
        \item Given that a customer does not renew, what is the probability they have only an auto policy?\vspace{40pt}
    \end{enumerate}\bigskip
    
    \item A corporate team has three managers, $M_1$, $M_2$ and $M_3$, who train 50\%, 30\% and 20\% of the employees, respectively. Further, 15\% of the employees trained by $M_1$ make mistakes on a certain process, 22\% for $M_2$ and 18\% for $M_3$. Find the probability that an employee makes a mistake.\vspace{100pt}
    
    \item Consider the game of Three. You shuffle a deck of three cards: 1, 2, 3. You draw cards without replacement until your total is 3 or more. You win if your total is 3.
    \item[] Let $C_i$ denote the event that card $C$ is the $i$th card drawn.
For example, $3_2$ is the event that the 3 was the second card drawn. Given that you win, find the probability that the Card 3 is drawn.\vspace{100pt}% Theory Lecture 4 example (page 7)
    
    \item Suppose we have two coins. We know one of the two coins is biased and comes up heads with probability 3/4, we will call this Coin 1. And we know the other coin is fair and comes up heads with probability 1/2, we will call this Coin 2. However, we don't actually know which coin is which. Suppose you pick a coin randomly and flip it.%Theory in-class 2
    \begin{enumerate}
        \item What is the probability that the outcome of the coin flip is a head? \\\textit{HINT: Think carefully about the stages of the tree diagram.}\vspace{100pt}
        \item Given that the outcome of the flip is a head, what is the probability that you picked up the biased coin?\vspace{40pt}
    \end{enumerate}\bigskip
    
    \item The probability that a randomly chosen make has a circulation problem is 0.3. Males who have a circulation problem are twice as likely to be smokers as those who do not have a circulation problem.%Actex 3-63 (with different numbers)
    \item[] What is the conditional probability that a male has a circulation problem, given that he is a smoker?\vspace{100pt}
    
\end{enumerate}

\end{document}