\documentclass{article}
\usepackage{style-assessments}

% define macros (/shortcuts)
\newcommand{\comp}{{\sim}}			% shortcut for tilde without extra space, using for complement

\begin{document}

\hspace{375pt}Name:

\begin{center}
{\Huge MATH 320: Review Part 1}

\end{center}

% problem types summary
% 1) 
% 2) 
% 3) 

\bigskip\bigskip

\begin{enumerate}
    \item Passwords for your iPhone require 6 characters.\bigskip
    \begin{enumerate}
        \item If the first three must be digits 0-9 and the last three must be lowercase letters a-z, how many different passwords can you make?\vspace{30pt}
        \item Using the same scenario from part (a), now suppose you cannot reuse digits / letters. How many different passwords can you make?\vspace{30pt}
        \item What is the probability your friend guesses your password correctly using the scenario from part (b).\vspace{30pt}
    \end{enumerate}\bigskip
    
    \item There is a group of 145 supporters representing England, Argentina or a different country for the world cup.\bigskip
    \begin{enumerate}
       \item Suppose 50 people support only England, 35 people support England and Argentina, and 20 support a different country. How many people support England or Argentina?\vspace{80pt}
       \item Now suppose 85 people support a different country. How many people support  England or Argentina?\vspace{80pt}
       \item Suppose 40 people support England. 75 people support Argentina and of those 25 support England. How many support a different country?\vspace{80pt}
   \end{enumerate}\bigskip
   
   \item There are two events: $A = $ watching Brazil vs Croatia and $B = $ watching Portugal vs Morocco. Let $P(A) = 0.6$, $P(B) = 0.35$ and $P(A \cap B) = 0.12$. Find the following:\bigskip
   \begin{enumerate}
       \item $P(A \mid B) = $\vspace{40pt}
       \item $P(B \mid A = $\vspace{40pt}
       \item $P(A \mid A \cup B) = $\vspace{40pt}
       \item $P(\comp B \mid \comp A) = $\vspace{40pt}
       \item Suppose now $A$ and $B$ are independent and we do not know $P(B)$ (we still know all other information in the setup). Find $P(B)$.\vspace{40pt}
   \end{enumerate}\bigskip
   
    \item There are 8 teams left in the world cup.\bigskip
    \begin{enumerate}
        \item How many different ways can the top 4 teams finish?\vspace{30pt}
        \item 4 teams move onto the next round, how many different ways can these 4 teams be selected?\vspace{30pt}
        \item What is the probability Morocco and Netherlands are selected to move onto the next round?\vspace{50pt}
        \item What is the probability Morocco or Netherlands are selected to move onto the next round?\vspace{50pt}
    \end{enumerate}\bigskip
    
    \item Suppose you roll an 8-sided die. Let $A =$ odd numbers and $B =$ numbers less than or equal to 4. Show if $A$ and $B$ are independent or not mathematically using TWO different ways.\vspace{100pt}
    
    \item France scores 0 goals in 20\% of their games, one goal in 50\% of their games, and 2 goals in 30\% of their games. When they do not score, there is a 20\% chance of winning; when they score one goal, there is a 40\% chance of winning; when they score two goals, there is a 80\% chance of winning. Assume they can only win or lose (no ties).\vspace{150pt}
    \begin{enumerate}
        \item What is the probability France wins?\vspace{70pt}
        \item What is the probability France loses?\vspace{70pt}
        \item If France wins, what is the probability that they scored one goal?\vspace{70pt}
        \item If France loses, what is the probability that they scored?\vspace{70pt}
    \end{enumerate}
\end{enumerate}

\end{document}