---
title: "HW 1 -- R STARTER"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    df-print: kable
execute: 
  warning: false
  message: false
---

## Simulation 1 

### a) Population distribution

Let $X \sim \text{Gamma}(\alpha = 4, \beta = 1/2)$.

```{r}


```

### b and c) Generate random samples and calculate statistic

Let $Y = \text{median}\{X_1, \ldots, X_{30}\}$.

```{r}


```

### d) Plot simulated sampling distribution of $Y = \text{median}\{X_1, \ldots, X_{30}\}$

```{r}


```

Sampling distribution seems to be centered at the population median.

### e) Calculate the mean and standard deviation of the simulated sampling distribution

Calculate $\hat{\mu}_Y$ and $\hat{\sigma}_Y$.

- We don't know the theoretical expected value and variance of the statistic, so can't compare the simulated results to anything.

```{r}


```

### f) Estimate the probability the sample statistic is within two standard deviations of its mean

Calculate $P(\hat{\mu}_Y - 2 \hat{\sigma}_Y < Y < \hat{\mu}_Y + 2 \hat{\sigma}_Y)$.

```{r}


```

## Simulation 2

### a) Population distribution

Let $X \sim \text{??}(??)$.

```{r}


```

### b and c) Generate random samples and calculate statistic

Let $Y = \text{??}\{X_1, \ldots, X_{30}\}$.

```{r}


```

### d) Plot simulated sampling distribution of $Y = \text{max}\{X_1, \ldots, X_{30}\}$

```{r}



```

### e) Calculate the mean and standard deviation of the simulated sampling distribution

Calculate $\hat{\mu}_Y$ and $\hat{\sigma}_Y$.

```{r}



```

### f) Estimate the probability the sample statistic is within two standard deviations of its mean

Calculate $P(\hat{\mu}_Y - 2 \hat{\sigma}_Y < Y < \hat{\mu}_Y + 2 \hat{\sigma}_Y)$.

```{r}



```