---
title: "In-Class 8"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    df-print: kable
execute: 
  warning: false
  message: false
---

```{r}

# load packages
library(tidyverse)

```

## Load data

The data we will be working with contains information about the housing market in Ames, TX.

We need to load the following RData file, which contains a training and testing dataset.

```{r}

# read in training and testing data
load('files/data/regression-data.RData')

# preview training data
glimpse(data_train)

```

## Intial EDA 

The model we want to create is: `sale_price ~ < set of Xs >`. Let's first inspect the response variable.

```{r}

# plot response variable
# -> seems skewed right and not normal
hist(data_train$sale_price)
qqnorm(data_train$sale_price)

# -> try log transformation and recheck
hist(log(data_train$sale_price))
qqnorm(log(data_train$sale_price))

```

The distribution of `sale_price` appears to be much more normal after transforming. So let's use `log(sale_price) ~ < set of Xs >` as our model.

This is technically a regression with a transformed response, so it is a generalized linear model (GLM). Once we make the transformation, everything works as usual, only the units and interpretations change.

## Variable selection

Now try to find some $X$ variables to use in a model.

```{r}



```

## Prediction

Once you have some candidate models, see which one is the best by calculating the $RMSE$.

```{r}



```