---
title: "R Notes 1-3 -- R Practice Session STARTER"
format:
  html:
    embed-resources: true
    toc: true
    toc-location: left
    df-print: kable
execute: 
  warning: false
  message: false
---

## Overview

Below are some basic problems to practice the concepts we have learned so far in R.


### Problem 1

Suppose we have the following discrete pmf for X:

| $x$ | $f(x)$ |
|-----|--------|
| 2   | 0.20   |
| 4   | 0.13   |
| 5   | 0.07   |
| 8   | 0.21   |
| 10  | 0.14   |
| 11  | 0.09   |
| 13  | 0.14   |

a)  Create two vectors, one for $x$ and one for $f(x)$.

```{r}


```

b)  Calculate the expected value of $X$ using your vectors.

```{r}


```

c)  Calculate $P(X > 7)$ by subsetting the $f(x)$ vector with a logical test.

```{r}


```

d)  Find the mode of $X$ by selecting the correct $x$ value based on a logic check of $f(x)$. NOTE: To check equality, use `==`.

```{r}


```

e)  Plot the pmf of $X$.

```{r}


```

### Problem 2

Generate a random sample from of size $n = 100$ from $\text{Gamma}(\text{shape } (\alpha) = 5, \text{rate } (\beta) = 2.5)$ and save it as a vector.

a)  Count the number of values in your random sample that are greater than 2 using `sum()` on a logical check of your vector.

```{r}


```

b)  Calculate the sample probability of less than 1.5 using `mean()` on a logical check of your vector.

```{r}


```

c)  Search how to and compute the min, max, median, IQR, and range of your random sample.

```{r}


```

d)  Create a boxplot of your random sample.

```{r}


```

e)  Search how to extract any outliers based on your boxplot extract.

```{r}


```
